---
title: "cosore-working-with-data"
author: "Ben Bond-Lamberty"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
%\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  ---
  
```{r setup, include = FALSE}
  knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>"
  )
```
  
# Introduction

Soil respiration--the flux of CO2 from the soil surface to the atmosphere--has been measured by continuous (typically hourly) measurement systems in many places over the last 25 years. The goal of the COSORE database is to collect many of these data into a single open community resource to support data synthesis and transparency. 

The database is distributed as an R package. Use the `devtools` package
to install COSORE: `devtools::install_github()`.

(For non-R users, the database is also distrubuted as a zip file of comma-separated files for non-R users, available through GitHub releases [here](https://github.com/bpbond/cosore/releases). When you download and extract the file, there's a README, the data in two separate formats, and a number of other files, including a version of this vignette.)

But how do we work with this database, exactly? Let's start by loading it into R:

```{r}
library(cosore)
```

The database is comprised of a collection of _datasets_, each converted to a standard format and units. A dataset is one or more files of continuous (automated) soil respiration data, with accompanying metadata, with all measurements taken at a single site and with constant treatment assignments.

In practice, the `COSORE` database can be thought of as a two-level nested list of data frames:

```
cosore
|- dataset 1
|   |- description table (a data.frame)
|   |- contributors table (ditto...)
|   |- ports table
|   |- columns table
|   |- ancillary table
|   |- data table
|   |- diagnostics table
|- dataset 2
|   |- description table
|   |- contributors table
|   |- ports table
|   |- columns table
|   |- ancillary table
|   |- data table
|   |- diagnostics table
|- dataset 3
|   |- etc.
```

For most analyses we want to extract one or more of these pieces and combine them--for example, to get a single dataset, a table of contributors, or an overview of the entire database. 

With this in place, we extract a combined `description` table for the entire database and take a look at it:

```{r description-table}
db_info <- csr_database()
tibble::glimpse(db_info)
```

Lots of information; note much of this is summarized in the `Report-all.html` file included with the data download.

# Exploring a single dataset

To begin, we pick a single dataset (`d20190415_VARNER`), get some information about it, and plot it.

```{r desc-varner}
varner <- csr_dataset("d20190415_VARNER")
tibble::glimpse(varner$description)
```

The `description` table gives the basic information about this dataset:
where it was measured, the time zone that the `data` timestamps are in, 
instrument used, and citation and acknowledgment information.

```{r sr-varner}
sr <- varner$data
nrow(sr)
summary(sr)
```

This dataset has `r nrow(sr)` observations; extends from April 2003 to December 2006; and soil respiration was measured on eight ports, along with air and 5 cm soil temperature. Visualizing it:

```{r visualize-varner}
library(ggplot2)
ggplot(sr, aes(CSR_TIMESTAMP_BEGIN, CSR_FLUX, color = CSR_PORT)) +
  geom_point(size = 0.5) + ggtitle("Flux over time")

library(lubridate, warn.conflicts = FALSE)
ggplot(sr, aes(CSR_T5, CSR_FLUX)) + 
  facet_wrap(~month(CSR_TIMESTAMP_BEGIN), scales = "free") +
  geom_point(size = 0.5, alpha = 0.5) +
  ggtitle("5 cm soil temp vs. flux, by month")
```

Did these eight different ports (chambers) represent different treatments?
We might want to exclude treatment collars, or color them differently 
in the plots above. The `ports` table holds this information.

```{r ports-varner}
varner$ports
```

From this we see that the only `CSR_PORT` entry is zero, and has a `CSR_TREATMENT` of "None", i.e. all ports/collars had the same treatment, and there _wasn't_ any treatment. Also, the collars were `r varner$ports$CSR_AREA` cm2.

Finally, we can use the `description` table information to get a full citation:

```{r pub-varner}
doi <- varner$description$CSR_PRIMARY_PUB
print(doi)
library(rcrossref)
try(cr_cn(dois = doi, format = "text"))
```

# Selecting and combining multiple datasets

Time for something more ambitious: let's examine how soil respiration varies over the course of the day in temperate deciduous forests.

```{r}
dbf_datasets <- subset(db_info, CSR_IGBP == "Deciduous broadleaf forest")$CSR_DATASET
tdf <- csr_table("description", dbf_datasets)
```


```{r dbf-map}
# Make a map of these datasets
library(sp)
library(leaflet)
map <- data.frame(lon = tdf$CSR_LONGITUDE, lat = tdf$CSR_LATITUDE)
coordinates(map) <- ~lon + lat
leaflet(map) %>% 
  addMarkers() %>% 
  addTiles()
```

There are `r nrow(tdf)` datasets here. Extract and visualize their data:

```{r dbf-data}
tdf_dat <- csr_table("data", dbf_datasets, quiet = TRUE)
ggplot(tdf_dat, aes(CSR_TIMESTAMP_BEGIN, CSR_FLUX, color = CSR_DATASET)) + 
  geom_point(size = 0.5) + 
  coord_cartesian(ylim = c(0, 20)) +
  ggtitle("Flux over time, colored by dataset")
```

The original question we were interested in was how respiration varies over the course of the day. Say we're also interested in site latitude as a possible covariate, requiring us to join together two of the tables we've extracted.

```{r dbf-day}
site_info <- tdf[c("CSR_DATASET", "CSR_SITE_NAME", "CSR_LATITUDE")]
# join two tables together
tdf_combined <- merge(tdf_dat, site_info, by = "CSR_DATASET")
# add some new fields
tdf_combined$Hour <- hour(tdf_combined$CSR_TIMESTAMP_BEGIN)
tdf_combined$Month <- hour(tdf_combined$CSR_TIMESTAMP_BEGIN)
tdf_combined$Year <- hour(tdf_combined$CSR_TIMESTAMP_BEGIN)

# for each month, compute mean flux for each hour of the day
tdf_smry <- aggregate(CSR_FLUX ~ CSR_DATASET + CSR_LATITUDE + Year + Month + Hour,
                      FUN = mean, data = tdf_combined)

ggplot(tdf_smry, aes(Hour, CSR_FLUX, color = CSR_LATITUDE)) + 
  geom_point(size = 0.5) + facet_wrap(~Month) +
  ggtitle("Flux by hour of day and month of year")
```

(Note that for simplicity, we're assuming above that data timestamps are in _local_ time, but that might not be true, e.g. if an instrument is set to UTC; see the `description` table.)

# Other tables and data

This vignette has mostly focused on the `description` and `data` tables, and briefly mentioned `ports`. Others include:

* The `contributors` table: dataset contributors. The first person listed should be considered the primary point of contact.
* `ancillary`: ancillary site-level data such as leaf area index, net primary production, soil texture, etc. All optional.

Two additional tables provide information about the processing of raw data into standardized COSORE datasets:

* `columns`: describes how columns in the original dataset (i.e. as contributed) were mapped to the COSORE standard fields, including any unit changes or transformations.
* `diagnostics`: metadata about the data ingestion process: rows and columns removed, errors, etc.

These can all be extracted using the `csr_table()` function shown above.

# Useful package functions for users

* `csr_database()`, demonstrated above, provides a database overview
* `csr_dataset()`, demonstrated above, loads a single dataset
* `csr_table()`, demonstrated above, load a single _table_ across multiple datasets
* `csr_report_database()` generates a HTML summary of the entire database
* `csr_report_dataset()` generates an HTML summary of a single dataset

# Feedback and contributions

Feedback is welcome on any aspects of the database design, strengths, limitations, formats, documentation...please [open a GitHub issue](https://github.com/bpbond/cosore/issues/new) or email; see the README.

Interested in contributing data to COSORE? Pleae [contact the maintainer](mailto:bondlamberty@pnnl.gov) or open a GitHub issue.
