---
title: "Working with COSORE data"
author: "Ben Bond-Lamberty"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with COSORE data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>"
)
```

Let's 

```r
cosore <- readRDS("cosore_date.RDS")
```

```{r, echo=FALSE}
# For the vignette, we just read from the drake cache, not use readRDS as above
cosore <- drake::readd("all")
```

The `cosore` data are structured as a two-level nested list of data frames:

```
cosore
 |- dataset 1
 |   |- description table (a data.frame)
 |   |- contributors table (same...)
 |   |- ports table
 |   |- columns table
 |   |- ancillary table
 |   |- data table
 |- dataset 2
 |   |- description table
 |   |- contributors table
 |   |- ports table
 |   |- columns table
 |   |- ancillary table
 |   |- data table
 |- dataset 3
 |   |- etc.
```

For most analyses we want to extract one or more of these pieces and combine them--for example, to get a single table of contributors. There are [various](https://cran.r-project.org/package=purrr) [packages](https://cran.r-project.org/package=rlist) for dealing with nested lists, but we can also write our own short extractor function:

```{r, echo=TRUE}
csr_table <- function(cosore, table_name) {

  extract <- function(x, table_name) {
    if(is.null(x[[table_name]])) { return(NULL) }

    x[[table_name]]$CSR_DATASET <- x$dataset_name
    x[[table_name]]
  }

  dplyr::bind_rows(lapply(cosore, extract, table_name = table_name))
}
```

With this in place, we extract a combined `description` table for the entire database and take a look at it:

```{r}
desc <- csr_table(cosore, "description")
tibble::glimpse(desc)
```



## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

