---
title: "COSORE dataset report"
date: "`r format(Sys.time(), '%d %B %Y')`"
params:
  dataset: NULL
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(ggplot2)
theme_set(theme_bw())
ds <- params$dataset
#library(drake)
#ds <- readd("dat_.d20190415_VARNER.")
```

# `r ds$dataset_name`


```{r}
d <- ds$description
```

## Site information

Info          |  Value
------------- | ----
Site          | `r d$Site_name`
Location      | `r d$Latitude` N, `r d$Longitude` E; `r d$Elevation` m
UTC offset    |  `r d$UTC_offset`
IGBP          | `r d$IGBP`

```{r map, echo=FALSE}
library(sp)
library(leaflet)
df <- data.frame(lon = ds$description$Longitude, lat = ds$description$Latitude)
coordinates(df) <- ~lon + lat
leaflet(df) %>% 
  addMarkers() %>% 
  addTiles(options = providerTileOptions(minZoom = 3, maxZoom = 5))
```

## Reference information

Info            |  Value
--------------- | ----
Publication (URL or DOI)  | `r d$Primary_pub`
Other publications        | `r d$Other_pubs`
Data (URL or DOI)         | `r d$Data_URL`

**Acknowledgment text**: `r d$Acknowledgment`

## Contributors

```{r, echo=FALSE}
knitr::kable(unlist(ds$contributors), col.names = c("Value"))
```

## Measurement information

**Instrument**: `r d$Instrument`

```{r, echo=FALSE}
knitr::kable(ds$ports)
knitr::kable(ds$columns)
```

## Data diagnostics

Info                |  Value
------------------- | ----
Records             | `r d$Records`
Records (error)     | `r sum(ds$data$CSR_ERROR)`
Records removed (no flux)   | `r d$Records_removed`
Records (flux < 0)  | `r sum(ds$data$CSR_FLUX < 0, na.rm = TRUE)`
Records (flux > 50) | `r sum(ds$data$CSR_FLUX > 50, na.rm = TRUE)`

**Flux summary**:

```{r diagnostics}
summary(ds$data$CSR_FLUX)
dsd <- subset(ds$data, !is.na(CSR_FLUX))
ggplot(dsd, aes(CSR_TIMESTAMP, CSR_FLUX, color = CSR_PORT)) + 
  geom_line() + facet_grid(CSR_PORT~.)
ggplot(dsd, aes(CSR_TIMESTAMP, log(CSR_FLUX), color = CSR_PORT)) +
  geom_point(size = 0.2) + 
  geom_hline(yintercept = log(50), linetype = 2)
ggplot(dsd, aes(x = CSR_FLUX)) + geom_histogram(bins = 30)
```
