---
title: "COSORE dataset report"
date: "`r format(Sys.time(), '%d %B %Y')`"
params:
  dataset: NULL
  quick: FALSE
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(ggplot2)
theme_set(theme_bw())
library(cosore)
library(lubridate)

ds <- params$dataset

# Compute a mean timestamp for ease of use below
dsd <- ds$data
dsd$CSR_TIMESTAMP <- do.call(c, Map(function(x, y) mean(c(x, y)), 
                                    ds$data$CSR_TIMESTAMP_BEGIN, 
                                    ds$data$CSR_TIMESTAMP_END))
quick <- params$quick
```

# `r ds$description$CSR_DATASET`

```{r, echo=FALSE}
d <- ds$description
diag <- ds$diagnostics

if(embargoed(ds)) {
  cat("---------------------------------------------------------------\n")
  cat("EMBARGO FLAG IS SET--DATA WILL NOT BE RELEASED\n")
  cat("---------------------------------------------------------------\n")
}
```

## Site information

Info          |  Value
------------- | ----
Site          | `r d$CSR_SITE_NAME`
Location      | `r d$CSR_LATITUDE` N, `r d$CSR_LONGITUDE` E
Elevation     | `r d$CSR_ELEVATION` m
IGBP          | `r d$CSR_IGBP`
Network       | `r paste(d$CSR_NETWORK, d$CSR_SITE_ID)`
Timezone      | `r d$CSR_TIMEZONE`
Notes         | `r d$CSR_NOTES`

```{r map, echo=FALSE}
library(sp)
library(leaflet)
desc <- ds$description
df <- data.frame(label = paste(desc$CSR_DATASET, desc$CSR_SITE_NAME),
                 lon = desc$CSR_LONGITUDE, lat = desc$CSR_LATITUDE)
coordinates(df) <- ~lon + lat
x <- addMarkers(leaflet(df), label = df$label)
addTiles(x, options = providerTileOptions(minZoom = 3, maxZoom = 5))
```

## Reference information

Info            |  Value
--------------- | ----
Publication (URL or DOI)  | `r d$CSR_PRIMARY_PUB`
Other publications        | `r d$CSR_OTHER_PUBS`
Data (URL or DOI)         | `r d$CSR_DATA_URL`

**Acknowledgment text**: `r d$CSR_ACKNOWLEDGMENT`

## Contributors

```{r contrib_table, echo=FALSE}
x <- knitr::kable(unlist(ds$contributors), format = "html", col.names = c("Value")) 
kableExtra::kable_styling(x, full_width = FALSE, position = "left")
```

## Measurement information

**Instrument**: `r d$CSR_INSTRUMENT`

**Measurement length (s)**: `r d$CSR_MSMT_LENGTH`

**File format**: `r d$CSR_FILE_FORMAT`

**Timestamp format**: `r d$CSR_TIMESTAMP_FORMAT`

**Timestamp timezone**: `r d$CSR_TIMESTAMP_TZ`

```{r msmt_table, echo=FALSE}
dsp <- ds$ports
x <- knitr::kable(dsp, format = "html")
kableExtra::kable_styling(x, full_width = FALSE, position = "left")
```

## Observation density

```{r}
if(is.data.frame(dsd)) {
  x <- cosore:::compute_interval(dsd)
  results <- list()
  for(yr in na.omit(unique(x$Year))) {
    d <- x[x$Year == yr,]
    results[[yr]] <- tibble(Year = yr,
                            `Interval (min)` = round(weighted.mean(d$Interval, d$N, na.rm = TRUE), 1))
  }
  results <- cosore:::rbind_list(results)
  kableExtra::kable_styling(knitr::kable(results), full_width = FALSE, position = "left")
}
```

## Column mapping

```{r mapping, echo=FALSE}
x <- knitr::kable(ds$columns, format = "html")
kableExtra::kable_styling(x,full_width = FALSE, position = "left")
```

## Data processing diagnostics

Info                |  Value
------------------- | ----
Records             | `r diag$CSR_RECORDS`
Records removed (error)     | `r diag$CSR_RECORDS_REMOVED_ERR`
Records removed (invalid timestamp)     | `r diag$CSR_RECORDS_REMOVED_TIMESTAMP`
Bad timestamp examples     | `r diag$CSR_EXAMPLE_BAD_TIMESTAMPS`
Records removed (no flux)   | `r diag$CSR_RECORDS_REMOVED_NA`
Size                | `r format(object.size(ds), "Mb")`

**Flux summary**:

```{r flux-graphs}
if(is.data.frame(dsd) & !quick) {
  has_ch4 <- any(!is.na(dsd$CSR_FLUX_CH4))
  has_co2 <- any(!is.na(dsd$CSR_FLUX_CO2))
  dsd$Month = month(dsd$CSR_TIMESTAMP_BEGIN)
  
  #dsd <- subset(dsd, !is.na(CSR_FLUX_CO2) & CSR_FLUX_CO2 > 0)
  dsd$CSR_PORT <- as.factor(dsd$CSR_PORT)
  
  print(summary(dsd$CSR_FLUX_CO2))
  print(summary(dsd$CSR_FLUX_CH4))
  
  # flux over time
  if(has_co2) {
    p1 <- ggplot(dsd, aes(CSR_TIMESTAMP, CSR_FLUX_CO2, color = CSR_PORT)) + 
      geom_line() + facet_grid(CSR_PORT~., scales = "free_y")
    print(p1)
    
    smry <- aggregate(CSR_FLUX_CO2 ~ CSR_PORT + Month, data = dsd, mean)
    p3 <- ggplot(smry, aes(Month, CSR_FLUX_CO2, color = CSR_PORT, group = CSR_PORT)) + 
      geom_line()
    print(p3)
  }
  
  if(has_ch4) {
    p1_ch4 <- ggplot(dsd, aes(CSR_TIMESTAMP, CSR_FLUX_CH4, color = CSR_PORT)) + 
      geom_line() + facet_grid(CSR_PORT~., scales = "free_y")
    print(p1_ch4)
    
    smry <- aggregate(CSR_FLUX_CH4 ~ CSR_PORT + Month, data = dsd, mean)
    p3_ch4 <- ggplot(smry, aes(Month, CSR_FLUX_CH4, color = CSR_PORT, group = CSR_PORT)) + 
      geom_line()
    print(p3_ch4)
  }
  
  # Histogram
  if(has_co2) {
    p4 <- ggplot(dsd, aes(x = CSR_FLUX_CO2)) + geom_histogram(bins = 30)
    print(p4)
  }
  
  if(has_ch4) {
    p4_ch4 <- ggplot(dsd, aes(x = CSR_FLUX_CH4)) + geom_histogram(bins = 30)
    print(p4_ch4)
  }
}
```

**Temperature sensitivity**:

```{r tsensitivity}
if(is.data.frame(dsd) & !quick) {
  tcols <- grep("^CSR_T(CHAMBER|[0-9]+)", names(dsd))
  if(length(tcols)) {
    
    if(has_co2) {
      results <- list()
      for(col in tcols) {
        x <- dsd[c("CSR_FLUX_CO2", "CSR_PORT")]
        x$Temperature <- unlist(dsd[,col])
        x$Which_temp <- names(dsd)[col]
        results[[as.character(col)]] <- x
      }
      results <- cosore:::rbind_list(results)
      
      p1 <- ggplot(results, aes(Temperature, CSR_FLUX_CO2, color = CSR_PORT)) + 
        geom_point(size = 0.2, alpha = I(0.25)) + 
        geom_smooth(method = "lm", se = FALSE) +
        facet_wrap(~Which_temp, scales = "free")
      print(p1)
    }
    
    if(has_ch4) {
      results <- list()
      for(col in tcols) {
        x <- dsd[c("CSR_FLUX_CH4", "CSR_PORT")]
        x$Temperature <- unlist(dsd[,col])
        x$Which_temp <- names(dsd)[col]
        results[[as.character(col)]] <- x
      }
      results <- cosore:::rbind_list(results)
      
      p1_ch4 <- ggplot(results, aes(Temperature, CSR_FLUX_CH4, color = CSR_PORT)) + 
        geom_point(size = 0.2, alpha = I(0.25)) + 
        geom_smooth(method = "lm", se = FALSE) +
        facet_wrap(~Which_temp, scales = "free")
      print(p1_ch4)
      
    }
  }
}
```

**Soil moisture sensitivity**:

```{r smsensitivity}
if(is.data.frame(ds$data) & !quick) {
  smcols <- grep("^CSR_SM[0-9]+", names(dsd))
  if(length(smcols)) {
    
    if(has_co2) {
      results <- list()
      for(col in smcols) {
        x <- dsd[c("CSR_FLUX_CO2", "CSR_PORT")]
        x$SM <- unlist(dsd[,col])
        x$Which_sm <- names(dsd)[col]
        results[[as.character(col)]] <- x
      }
      results <- cosore:::rbind_list(results)
      p1 <- ggplot(results, aes(SM, CSR_FLUX_CO2, color = CSR_PORT)) + 
        geom_point(size = 0.2, alpha = I(0.25)) + 
        geom_smooth(method = "lm", se = FALSE) +
        facet_wrap(~Which_sm, scales = "free") +
        xlab("Soil moisture")
      print(p1)
    }
    
    if(has_ch4) {
      results <- list()
      for(col in smcols) {
        x <- dsd[c("CSR_FLUX_CH4", "CSR_PORT")]
        x$SM <- unlist(dsd[,col])
        x$Which_sm <- names(dsd)[col]
        results[[as.character(col)]] <- x
      }
      results <- cosore:::rbind_list(results)
      
      p1_ch4 <- ggplot(results, aes(SM, CSR_FLUX_CH4, color = CSR_PORT)) + 
        geom_point(size = 0.2, alpha = I(0.25)) + 
        geom_smooth(method = "lm", se = FALSE) +
        facet_wrap(~Which_sm, scales = "free") +
        xlab("Soil moisture")
      print(p1_ch4)
    }  
  }
}
```

## Ancillary data

Ancillary dataset has `r nrow(ds$ancillary)` rows.

```{r ancillary}
names(ds$ancillary)
```

