---
title: "COSORE dataset report"
date: "`r format(Sys.time(), '%d %B %Y')`"
params:
  dataset: NULL
  quick: FALSE
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(ggplot2)
theme_set(theme_bw())
library(cosore)
library(lubridate)

ds <- params$dataset

# Compute a mean timestamp for ease of use below
dsd <- ds$data
dsd$CSR_TIMESTAMP <- do.call(c, Map(function(x, y) mean(c(x, y)), 
                                        ds$data$CSR_TIMESTAMP_BEGIN, 
                                        ds$data$CSR_TIMESTAMP_END))
quick <- params$quick
```

# `r ds$description$CSR_DATASET`

```{r, echo=FALSE}
d <- ds$description
diag <- ds$diagnostics

if(!is.na(d$CSR_EMBARGO)) {
  cat("---------------------------------------------------------------\n")
  cat("EMBARGO FLAG IS SET--DATA WILL NOT BE RELEASED\n")
  cat("---------------------------------------------------------------\n")
}
```

## Site information

Info          |  Value
------------- | ----
Site          | `r d$CSR_SITE_NAME`
Location      | `r d$CSR_LATITUDE` N, `r d$CSR_LONGITUDE` E
Elevation     | `r d$CSR_ELEVATION` m
IGBP          | `r d$CSR_IGBP`
Network       | `r paste(d$CSR_NETWORK, d$CSR_SITE_ID)`
Timezone      | `r d$CSR_TIMEZONE`
Notes         | `r d$CSR_NOTES`

```{r map, echo=FALSE}
library(sp)
library(leaflet)
df <- data.frame(lon = ds$description$CSR_LONGITUDE, lat = ds$description$CSR_LATITUDE)
coordinates(df) <- ~lon + lat
x <- addMarkers(leaflet(df))
addTiles(x, options = providerTileOptions(minZoom = 3, maxZoom = 5))
```

## Reference information

Info            |  Value
--------------- | ----
Publication (URL or DOI)  | `r d$CSR_PRIMARY_PUB`
Other publications        | `r d$CSR_OTHER_PUBS`
Data (URL or DOI)         | `r d$CSR_DATA_URL`

**Acknowledgment text**: `r d$CSR_ACKNOWLEDGMENT`

## Contributors

```{r contrib_table, echo=FALSE}
x <- knitr::kable(unlist(ds$contributors), format = "html", col.names = c("Value")) 
kableExtra::kable_styling(x, full_width = FALSE, position = "left")
```

## Measurement information

**Instrument**: `r d$CSR_INSTRUMENT`

**Measurement length (s)**: `r d$CSR_MSMT_LENGTH`

**File format**: `r d$CSR_FILE_FORMAT`

**Timestamp format**: `r d$CSR_TIMESTAMP_FORMAT`

**Timestamp timezone**: `r d$CSR_TIMESTAMP_TZ`

```{r msmt_table, echo=FALSE}
dsp <- ds$ports
x <- knitr::kable(dsp, format = "html")
kableExtra::kable_styling(x, full_width = FALSE, position = "left")
```

## Observation density

```{r}
if(is.data.frame(dsd)) {
  x <- cosore:::compute_interval(dsd)
  results <- list()
  for(yr in na.omit(unique(x$Year))) {
    d <- x[x$Year == yr,]
    results[[yr]] <- tibble(Year = yr,
                            `Interval (min)` = round(weighted.mean(d$Interval, d$N, na.rm = TRUE), 1))
  }
  results <- cosore:::rbind_list(results)
  kableExtra::kable_styling(knitr::kable(results), full_width = FALSE, position = "left")
}
```

## Column mapping

```{r mapping, echo=FALSE}
x <- knitr::kable(ds$columns, format = "html")
kableExtra::kable_styling(x,full_width = FALSE, position = "left")
```

## Data processing diagnostics

Info                |  Value
------------------- | ----
Records             | `r diag$CSR_RECORDS`
Records removed (error)     | `r diag$CSR_RECORDS_REMOVED_ERR`
Records removed (invalid timestamp)     | `r diag$CSR_RECORDS_REMOVED_TIMESTAMP`
Bad timestamp examples     | `r diag$CSR_EXAMPLE_BAD_TIMESTAMPS`
Records removed (no flux)   | `r diag$CSR_RECORDS_REMOVED_NA`
Records (flux < `r diag$CSR_FLUX_LOW_LIM_CO2`)  | `r diag$CSR_REMOVED_LOW_CO2`
Records (flux > `r diag$CSR_FLUX_HIGH_LIM_CO2`) | `r diag$CSR_REMOVED_HIGH_CO2`
Bad temperature values removed | `r diag$CSR_BAD_TEMPERATURE`
Size                | `r format(object.size(ds), "Mb")`

**Flux summary**:

```{r flux-graphs}
if(is.data.frame(dsd) & !quick) {
  dsd <- subset(dsd, !is.na(CSR_FLUX_CO2) & CSR_FLUX_CO2 > 0)
  dsd$CSR_PORT <- as.factor(dsd$CSR_PORT)
  
  print(summary(ds$data$CSR_FLUX_CO2))
  
  # flux over time
  p1 <- ggplot(dsd, aes(CSR_TIMESTAMP, CSR_FLUX_CO2, color = CSR_PORT)) + 
    geom_line() + facet_grid(CSR_PORT~., scales = "free_y")
  print(p1)
  
  p2 <- ggplot(dsd, aes(CSR_TIMESTAMP, CSR_FLUX_CO2, color = CSR_PORT)) +
    geom_point(size = 0.2) + scale_y_log10()
  print(p2)
  
  dsd$Month = month(dsd$CSR_TIMESTAMP)
  smry <- aggregate(CSR_FLUX_CO2 ~ CSR_PORT + Month, data = dsd, mean)
  ggplot(smry, aes(Month, CSR_FLUX_CO2, color = CSR_PORT, group = CSR_PORT)) + 
    geom_line() ->
    p3
  print(p3)
  # histogram
  p4 <- ggplot(dsd, aes(x = CSR_FLUX_CO2)) + geom_histogram(bins = 30)
  print(p4)
}
```

**Temperature sensitivity**:

```{r tsensitivity}
if(is.data.frame(dsd) & !quick) {
  tcols <- grep("^CSR_T(CHAMBER|[0-9]+)", names(dsd))
  if(length(tcols)) {
    results <- list()
    for(col in tcols) {
      x <- dsd[c("CSR_FLUX_CO2", "CSR_PORT")]
      x$Temperature <- unlist(dsd[,col])
      x$Which_temp <- names(dsd)[col]
      results[[as.character(col)]] <- x
    }
    results <- cosore:::rbind_list(results)
    
    p1 <- ggplot(results, aes(Temperature, CSR_FLUX_CO2, color = CSR_PORT)) + 
      geom_point(size = 0.2, alpha = I(0.25)) + 
      geom_smooth(method = "lm", se = FALSE) +
      facet_wrap(~Which_temp, scales = "free")
    print(p1)
  }
}
```

**Soil moisture sensitivity**:

```{r smsensitivity}
if(is.data.frame(ds$data) & !quick) {
  smcols <- grep("^CSR_SM[0-9]+", names(dsd))
  if(length(smcols)) {
    results <- list()
    for(col in smcols) {
      x <- dsd[c("CSR_FLUX_CO2", "CSR_PORT")]
      x$SM <- unlist(dsd[,col])
      x$Which_sm <- names(dsd)[col]
      results[[as.character(col)]] <- x
    }
    results <- cosore:::rbind_list(results)
    
    p1 <- ggplot(results, aes(SM, CSR_FLUX_CO2, color = CSR_PORT)) + 
      geom_point(size = 0.2, alpha = I(0.25)) + 
      geom_smooth(method = "lm", se = FALSE) +
      facet_wrap(~Which_sm, scales = "free") +
      xlab("Soil moisture")
    print(p1)
  }
}
```
